#Test link cpp and R

This script tries to link the cpp and R file to access simulated distribution output in R.

load data of the simulation all scenarios

```{r warning=FALSE}
scenarios = c()
data = data.frame()
getwd()
baseDir <- "../distribution_cpp/output/fortes_lima_nStep=4097"

#Loop over all scenarios folders
for (scenarioDir in list.dirs(baseDir, recursive = FALSE)) {
  scenario = basename(scenarioDir)
  scenarios = append(scenarios,scenario)
  #Extract the data from simulations
  filename <- paste(baseDir,"/",scenario,"/",scenario,"sumStat.txt", sep='')
  dat <- read.table(filename, header = F, sep = ';', dec = ".")
  
  data = rbind(data, dat)
}
ncol(data)
```

Add labels and combine the different models simulations to train abcrf from Verdu.

```{r}
data$model = rep(scenarios, each = 10000)
head(data)
```

```{r}
# Plot confusion matrix
library(ggplot2)
library(reshape2)
library(svglite)
library(abcrf)
library(dplyr)

# Split data into model identifiers (first column) and summary statistics
model_identifiers <- as.factor(data[, 18])  # Convert model IDs to a factor
summary_stats <- data[, -c(1,18)] #Remove labels and id and min (because min not enough different between models)

#Merge into a new dataframe and keep only rows without NA and Inf
data = data.frame(model_identifiers, summary_stats)
finalData = data[complete.cases(data),]
finalData <- finalData[is.finite(rowSums(finalData[, -c(1)])),]

#Check the distribution of the removed rows among the scenarios
finalData %>% count(model_identifiers)
print(finalData)

#Fit the model
fit <- abcrf(model_identifiers ~ ., data = finalData, ntree=1000, paral=TRUE, ncores=8, sampsize=8000)

#Print a summary of the trained model
print(fit)

#Compute OOB errors wrt number of trees
OOB_errors = err.abcrf(object = fit, training = finalData, paral=TRUE, ncores = 8)

#Make the confusion matrix
conf_matrix <- fit$model.rf$confusion.matrix
conf_matrix_df <- as.data.frame(as.table(conf_matrix))

#Compute percentage of errors
conf_matrix_df <- conf_matrix_df[conf_matrix_df$Var2!="class.error",] %>%
  group_by(Var1) %>%
  mutate(Percentage = round(Freq / sum(Freq)*100, digits=1))

#Order  of display
scenario_order <- c("Afr2P-Eur2P", "AfrDE-Eur2P", "AfrIN-Eur2P", "Afr2P-EurDE", "AfrDE-EurDE", "AfrIN-EurDE", "Afr2P-EurIN", "AfrDE-EurIN", "AfrIN-EurIN")

#Plot the confusion matrix
conf_mat = ggplot(conf_matrix_df, aes(x = Var1, y = Var2, fill = Percentage)) +
  geom_tile() +
  geom_text(aes(label = Percentage), color = "black", size = 4) +
  scale_fill_gradient(limits = c(0,100), low = "white", high = "darkolivegreen3") +
  scale_x_discrete(limits = scenario_order) +
  scale_y_discrete(limits = scenario_order) + 
  labs(x = "True model",
       y = "ABC-RF Predicted model",
       fill = "Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1), axis.title = element_text(size = 16)) 

conf_mat
ggsave(file="conf_mat.png", plot=conf_mat, width = 10, height = 5, device='png', dpi=390)
```


```{r}
# Load test to find simulations
filename_ASW = "../../Data/SummaryStats/final_sumstats_ASW.txt"
filename_ACB = "../../Data/SummaryStats/final_sumstats_ACB.txt"
data_obs_ASW = read.table(filename_ASW, header = T, sep = ' ', dec = ".")
data_obs_ACB = read.table(filename_ACB, header = T, sep = ' ', dec = ".")

print(data_obs_ASW[,c(29:44)]) #16 sum stat of interest
```


```{r}
#Predict the scenario
data_obs_ordered_ASW = data_obs_ASW[,c(29,30,31,32,33,34,44,35,36,37,38,39,40,41,42,43)]
data_obs_ordered_ACB = data_obs_ACB[,c(29,30,31,32,33,34,44,35,36,37,38,39,40,41,42,43)]
colnames(data_obs_ordered_ASW) = c("V2","V3","V4","V5","V6","V7","V8","V9","V10","V11","V12","V13","V14","V15","V16","V17")
colnames(data_obs_ordered_ACB) = c("V2","V3","V4","V5","V6","V7","V8","V9","V10","V11","V12","V13","V14","V15","V16","V17")

predicted_params_ASW <- predict(fit, obs = data_obs_ordered_ASW, training = finalData, paral=TRUE, ncores = 8)
predicted_params_ACB <- predict(fit, obs = data_obs_ordered_ACB, training = finalData, paral=TRUE, ncores = 8)
print(predicted_params_ASW)
print(predicted_params_ACB)

#Make a long table and change colnames
predicted_params_long_ASW = melt(predicted_params_ASW$vote)[,-1]
colnames(predicted_params_long_ASW) <- c("model", "votes")

predicted_params_long_ACB = melt(predicted_params_ACB$vote)[,-1]
colnames(predicted_params_long_ACB) <- c("model", "votes")

# Plot
prediction_fl_ASW = ggplot(predicted_params_long_ASW, aes(x = model, y = votes, fill = model)) +
  geom_bar(stat = "identity") +
  labs(x = "Model", y = "Number of Votes", title = "Votes per Scenarios") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

prediction_fl_ACB = ggplot(predicted_params_long_ACB, aes(x = model, y = votes, fill = model)) +
  geom_bar(stat = "identity") +
  labs(x = "Model", y = "Number of Votes", title = "Votes per Scenarios") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ggsave(file="vote_ASW.png", plot=prediction_fl_ASW, width = 10, height = 5, device='png', dpi=390)
ggsave(file="vote_ACB.png", plot=prediction_fl_ACB, width = 10, height = 5, device='png', dpi=390)
```


Here we plot the abcrf to verify the condition of application
```{r}
ASW_plot_LDA = plot(fit, training =finalData, obs = data_obs_ordered_ASW)
ACB_plot_LDA = plot(fit, training =finalData, obs = data_obs_ordered_ACB)

ASW_plot_LDA
ACB_plot_LDA
```
# Inference with selection

```{r warning=FALSE}
scenarios = c()
data_sel = data.frame()
getwd()
baseDir <- "../distribution_cpp/output/Selection_nStep=4097"

#Loop over all scenarios folders
for (scenarioDir in list.dirs(baseDir, recursive = FALSE)) {
  scenario = basename(scenarioDir)
  scenarios = append(scenarios,scenario)
  #Extract the data from simulations
  filename <- paste(baseDir,"/",scenario,"/",scenario,"sumStat.txt", sep='')
  dat <- read.table(filename, header = F, sep = ';', dec = ".")
  
  data_sel = rbind(data_sel, dat)
}
print(data_sel)
```

```{r}
# Plot confusion matrix
library(ggplot2)
library(reshape2)
library(svglite)
library(abcrf)
library(dplyr)

#add column of scenario
data_sel$model = rep(scenarios, each = 10000)

# Split data into model identifiers (first column) and summary statistics
model_identifiers <- as.factor(data_sel[, 18])  # Convert model IDs to a factor
summary_stats <- data_sel[, -c(1,7,8,18)] #Remove labels and id and min (because min not enough different between models)

#Merge into a new dataframe and keep only rows without NA and Inf
data_sel = data.frame(model_identifiers, summary_stats)
finalData = data_sel[complete.cases(data_sel),]
finalData <- finalData[is.finite(rowSums(finalData[, -c(1)])),]

#Check the distribution of the removed rows among the scenarios
finalData %>% count(model_identifiers)

#Fit the model
fit_sel <- abcrf(model_identifiers ~ ., data = finalData, ntree=1000, paral=TRUE, ncores=8, sampsize=8000)

#Print a summary of the trained model
print(fit_sel)

#Compute OOB errors wrt number of trees
#OOB_errors = err.abcrf(object = fit, training = finalData, paral=TRUE, ncores = 8)

#Make the confusion matrix
conf_matrix_sel <- fit_sel$model.rf$confusion.matrix
conf_matrix_df_sel <- as.data.frame(as.table(conf_matrix_sel))

#Compute percentage of errors
conf_matrix_df_sel <- conf_matrix_df_sel[conf_matrix_df_sel$Var2!="class.error",] %>%
  group_by(Var1) %>%
  mutate(Percentage = round(Freq / sum(Freq)*100, digits=1))

#Plot the confusion matrix
conf_mat_sel = ggplot(conf_matrix_df_sel, aes(x = Var1, y = Var2, fill = Percentage)) +
  geom_tile() +
  geom_text(aes(label = Percentage), color = "black", size = 4) +
  scale_fill_gradient(limits = c(0,100), low = "white", high = "darkolivegreen3") +
  labs(x = "True model",
       y = "ABC-RF Predicted model",
       fill = "Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 10, hjust = 1), axis.title = element_text(size = 16)) 

conf_mat_sel
ggsave(file="conf_mat_sel.png", plot=conf_mat_sel, width = 10, height = 5, device='png', dpi=390)
```

Inference of the scenario with selection

```{r}
#Predict the scenario
data_obs_ordered_ASW_sel = data_obs_ASW[,c(29,30,31,32,33,35,36,37,38,39,40,41,42,43)]
data_obs_ordered_ACB_sel = data_obs_ACB[,c(29,30,31,32,33,35,36,37,38,39,40,41,42,43)]
colnames(data_obs_ordered_ASW_sel) = c("V2","V3","V4","V5","V6","V9","V10","V11","V12","V13","V14","V15","V16","V17")
colnames(data_obs_ordered_ACB_sel) = c("V2","V3","V4","V5","V6","V9","V10","V11","V12","V13","V14","V15","V16","V17")

predicted_params_ASW_sel <- predict(fit_sel, obs = data_obs_ordered_ASW_sel, training = finalData, paral=TRUE, ncores = 8)
predicted_params_ACB_sel <- predict(fit_sel, obs = data_obs_ordered_ACB_sel, training = finalData, paral=TRUE, ncores = 8)
print(predicted_params_ASW_sel)
print(predicted_params_ACB_sel)

#Make a long table and change colnames
predicted_params_long_ASW_sel = melt(predicted_params_ASW_sel$vote)[,-1]
colnames(predicted_params_long_ASW_sel) <- c("model", "votes")

predicted_params_long_ACB_sel = melt(predicted_params_ACB_sel$vote)[,-1]
colnames(predicted_params_long_ACB_sel) <- c("model", "votes")

#Plots
prediction_ASW_sel = ggplot(predicted_params_long_ASW_sel, aes(x = model, y = votes, fill = model)) +
  geom_bar(stat = "identity") +
  labs(x = "Model", y = "Number of Votes", title = "Votes per Scenarios") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

prediction_ACB_sel = ggplot(predicted_params_long_ACB_sel, aes(x = model, y = votes, fill = model)) +
  geom_bar(stat = "identity") +
  labs(x = "Model", y = "Number of Votes", title = "Votes per Scenarios") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ggsave(file="vote_ASW_sel.png", plot=prediction_ASW_sel, width = 10, height = 5, device='png', dpi=390)
ggsave(file="vote_ACB_sel.png", plot=prediction_ACB_sel, width = 10, height = 5, device='png', dpi=390)
```


Here we plot the abcrf to verify the condition of application
```{r}
ASW_plot_LDA_sel = plot(fit_sel, training =finalData, obs = data_obs_ordered_ASW_sel)
ACB_plot_LDA_sel = plot(fit_sel, training =finalData, obs = data_obs_ordered_ACB_sel)

ASW_plot_LDA_sel
ACB_plot_LDA_sel

```
